---
title: "AddFactor"
author: "Rex Macey"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AddFactor}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = ""
)
```

This vignette introduces the `addFactor` function which is used to manipulate a set of factors.  The RAFI factors have some high correlations.  There are also other series we might want to add to the set of factors.  The `addFactor` function helps manipulate an original table of factor data (e.g. the RAFI data) into one we want to use to build models.

Specifically, we may want to:    
* Add factors representing the U.S., International Developed, and Emerging regions. These would be market-cap weighted indices.    
* Subtract these market-cap indices from the existing factors to reduce multi-collinearity.    

Examples of these operations are showin in this vignette.  We will use the dataset representing U.S. factors.

```{r, warning=FALSE, comment="", message=FALSE}
library(dplyr, quietly = TRUE)
library(factorModel, quietly = TRUE)
library(lubridate, quietly = TRUE)
data("factorReturns.US")
data("factorGroups")
head(factorReturns.US)
```

The following table shows the correlations among the factors.  Notice the high correlation between the factors.

```{r}
plot_corr(cor(factorReturns.US[,2:ncol(factorReturns.US)]))
```


First we will add a factor representing the Russell 2000 index using the iShares ETF (IWM). After we do that, we will add one for the US market represented by the iShares S&P 500 Index ETF (IVV) and subtract those returns from the other factors. We download prices and calculate monthly returns using the `get_monthly_returns` function.

**Get returns for the two ETFs.**
```{r}
sdate <- min(factorReturns.US$date) %m-% months(1)
edate <- max(factorReturns.US$date)
symbols <- c("IWM", "IVV")
monthly_returns <- get_monthly_returns(symbols = symbols, start_date = sdate, 
                                       end_date = edate)
```

The `get_monthly_returns` function created a 3 column table with the symbol, month (date) and Ra the monthly return.  Here are the first and last few rows of the table.

```{r comment=""}
head(monthly_returns)
tail(monthly_returns)
```


**Add the Russell 2000 (IWM) ETF series creating a new table.**  The new table will have the maximum common date range between the factor table and the series being added.
```{r}
factorUS <- AddFactor(factorReturns.US, 
                      monthly_returns %>% filter(symbol == "IWM") %>% 
                        select(c("date", "Ra")),
                      calcExcess = FALSE)
colnames(factorUS)[ncol(factorUS)] <- "USR2k"
head(factorUS)
```

**Add the S&P 500 (IVV) ETF series while subtracting that return from the other series.**

```{r}
factorUS <- AddFactor(factorUS, 
                      monthly_returns %>% filter(symbol == "IVV") %>% 
                        select(c("date", "Ra")),
                      calcExcess = TRUE)
colnames(factorUS)[ncol(factorUS)] <- "USMkt"
head(factorUS)
```

Comparing the correlation tables below, we can see the correlations are generally lower except for the two size factors (USSize and USR2k) which makes sense.  We wouldn't want both in a regression. 

```{r, echo=FALSE}
common_date_range <- FindMaxCommonDateRange(factorReturns.US$date, factorUS$date)
cortbl1 <- cor(factorReturns.US %>% filter(date >= common_date_range["start"] & 
                    date <= common_date_range["end"]) %>%
             select(2:ncol(factorReturns.US)))
cortbl2 <- cor(factorUS %>% filter(date >= common_date_range["start"] & 
                    date <= common_date_range["end"]) %>%
             select(2:ncol(factorUS)))
    
plot_corr(cortbl1)
plot_corr(cortbl2)
```

