---
title: "Description of Factor Return Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Description of Factor Return Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(factorModel)
data("factorReturns", package = "factorModel")
data("factorReturns.Intl", package = "factorModel")
data("factorReturns.US", package = "factorModel")
data("factorReturns.EM", package = "factorModel")
data("factorGroups", package = "factorModel")
```

## Overview

This vignette describes the data included with the factorModel package.  
Here are the datasets included with the package.  

```{r listOfDatasets, echo=FALSE, comment=""}
datasetlist <- (data(package="factorModel"))
datasetlist$results[, c("Item", "Title")]
```

* factorReturns are monthly returns for `r ncol(factorReturns)` factors across 3 regions: U.S.; International Developed; and Emerging.  
* factorReturns.US, factorReturns.Intl, and factorReturns.EM are subsets of factorReturns including only the factors for the regions.
* factorGroups is a list showing which factors are in which region.

Below are the returns for a few months for each region's factors.

```{r exampleData, echo=FALSE, comment=""}
cat("U.S.\n")
head(factorReturns.US)
tail(factorReturns.US)

cat("\n\nInternational Developed\n")
head(factorReturns.Intl)
tail(factorReturns.Intl)

cat("\n\nEmerging\n")
head(factorReturns.EM)
tail(factorReturns.EM)

cat("\n\nFactor Groups\n")
factorGroups
```

## Summary of Data

Below is a summary of the distribution of each factor.  Note that the three regions have different lengths of data series.  There are no obvious outliers.

```{r dataSummary, echo=FALSE, comment=""}
summary(factorReturns)
```

## Correlations

The following tables show the correlations.  We worry about correlations above 0.7 causing multi-collinearity in regressions.

```{r dataCorrelations, echo=FALSE, comment=""}
cat("U.S.\n")
plot_corr(cor(factorReturns.US[,2:ncol(factorReturns.US)]))
cat("\n\nInternational\n")
plot_corr(cor(factorReturns.Intl[,2:ncol(factorReturns.Intl)]))
cat("\n\nEmerging\n")
plot_corr(cor(factorReturns.EM[,2:ncol(factorReturns.EM)]))
```

